Chasing \rainbows{} with the Habitable Worlds Observatory

### 2023-03-09 src/tex/ms.tex
[showyourwork] overleaf sync

\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
%\textbf{Why glint [Kim]}
% probe of surface conditions; habitability
\cite{2001Natur.412..885F} first noted the potential for time-resolved photometry and polarimetry to identify water oceans on the surface of an exoplanet. \textcolor{blue}{A sentence about rotational mapping here.} \cite{2006astro.ph.10518M} further quantified how photometric and polarimetric scattered light phase variations might betray the presence of oceans on terrestrial exoplanets.   
%--> No it was talked about previous to this, per the paper cited here in fact... "Stam & Hovenier(2006) independently examined the observability of the polarized signatures of an Earth-likeextrasolar planet, including Rayleigh scattering and sea-surface glint. Schneider (2006) had similar ideas." -KB
\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets. They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases. \cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves. Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization. \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
\citet{Zugger_2011} note that ocean glint might be observed in near infrared opacity windows, because of the reduced Rayleigh scattering at long wavelengths. This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}. \cite{2019A&A...626A.129T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds. \cite{2022A&A...664A.172T}...

>>>>

%\textcolor{blue}{Kim will enhance the polarimetry stuff, and we should break into multiple paragraphs.}\\
%\textbf{Why glint [Kim]}
% probe of surface conditions; habitability
\cite{2001Natur.412..885F} first noted the potential for time-resolved photometry and polarimetry to identify water oceans on the surface of an exoplanet. Rotational variability of an exoplanet's reflected light colours can be inverted into a rough map of the planet's surface, with dark regions being suggestive of oceans \citep{2009ApJ...700..915C}, but specular reflection, a.k.a. glint, remains the most direct means to infer liquids on the surface of other worlds \citep[e.g., methane lakes on Titan;][]{2010GeoRL..37.7104S}.  \cite{2006astro.ph.10518M}  quantified how photometric and polarimetric scattered light orbital phase variations might betray the presence of oceans on terrestrial exoplanets.   
%--> No it was talked about previous to this, per the paper cited here in fact... "Stam & Hovenier(2006) independently examined the observability of the polarized signatures of an Earth-likeextrasolar planet, including Rayleigh scattering and sea-surface glint. Schneider (2006) had similar ideas." -KB
\cite{2008Icar..195..927W} simulated photometric and polarimetric phase curves of directly imaged terrestrial planets. They noted that surface oceans would be detectable because they are more reflective and exhibit a peak in polarization at crescent phases.

In practice, the clouds complicate the detection of ocean glint because they obscure the underlying surface and they are themselves non-Lambertian scatterers. \cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves. Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization. \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
\citet{Zugger_2011} note that ocean glint might be observed in near infrared opacity windows, because of the reduced Rayleigh scattering at long wavelengths. This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}. \cite{2019A&A...626A.129T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds. \cite{2022A&A...664A.172T}...
